Appendix part 2

### 2025-11-06 Output/authoritarian_only_pdf_pop_rev.Rmd
Near-final edits to lit review; removing redundant text and adjusting structure

---
output:
  word_document:
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
abstract: "Two models of the relationship between election manipulation and protest can be found in the literature. One approach holds that election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections. An alternative approach contends that election manipulation can be a sign of underlying regime strengths; under this framework, protest risk is less likely to deter manipulation. This article interrogates the assumptions underlying these models, and conducts a novel test of their implications. Using data on the duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly. This supports the ‘regime-strength’ model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding. It suggests researchers should look beyond protest risk in models of electoral manipulation."
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
  - \usepackage{multirow}
  - \usepackage{lscape}
  - \usepackage{pdflscape}
  - \usepackage{longtable}

@@

At first glance, it would appear straightforward to test these competing models. The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest @Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a. A low total vote-share amid severe manipulation is the strongest possible signal of incumbent weakness; it has had to rely on the 'cheap talk' of manipulation to produce paltry results, indicating that the incumbent's genuine popularity is very low and its likelihood of defeating the protest movement is likewise small. By contrast, the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant). Under the regime-strength model, the most favorable  election outcome for protesters is one where the incumbent's vote-share and manipulation level are low, a combination which indicate low natural popularity and diminished resource capacity.

@@

@@

As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest. This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength. Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power. As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength. Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among elite groups, the military and security services, or help from foreign patrons. These latter factors may be weakly correlated, if at all, with the incumbent's genuine popularity. This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle. This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012].

@@

The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely tied to popularity. At worst, election manipulation reveals low popularity (i.e. weakness). As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent. For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p. 1153). In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud---a pro-democracy outcome in which stronger opposition groups deter manipulation by incumbents. Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].

@@

@@

The second assumption in the regime-weakness framework concerns the nature of opposition protest. The framework is tightly linked to a view of electoral protest that is a centralized decision by rational actors pursuing narrowly strategic goals. As a result, they predict that the likelihood of protest should be highest when incumbents use massive fraud but still produce a low incumbent vote-share. This signal of weakness, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent. Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b]. As Beaulieu [-@Beaulieu2014a, p. 48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits."

Most fundamentally, regime-weakness models assume that opposition protest is *elite-led and instrumental*. The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure. Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.

Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006a]. Opposition elites, in turn, might seek to initiate protest to take advantage of this popular discontent for strategic, rather than instrumental, reasons. They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010]. They may stage protests in order to identify new participants, forge new social ties [@barberaModelProtestsRevolution2019] and test new tactics. They may seek to communicate information about the government's unpopularity, to set the stage for larger, future protests [@lohmannDynamicsInformationalCascades1994]. Protests may be staged as part of internal factional rivalries within the opposition [@Buckles2017], or to drive media coverage of issues important to the opposition [@huangMediaProtestDiffusion2019;@jenningsStreetDemonstrationsMedia2019]. Indeed, opposition elites may take advantage of the bare fact of an election to challenge the incumbent, irrespective of the level of manipulation or the incumbent's vote-share [@Pop-Eleches2015].

@@

## Election manipulation and regime weakness

<!-- The central feature of these models is that protest decisions are generally decentralized, and made with less consideration for the likelihood that protests will overthrow the incumbent. Belief in a stolen election---one the incumbent would have lost in the absence of manipulation---is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007;@kuntz_more_2009]. These models thus tend to predict that protests are most likely when margins of victory are low and election manipulation high, suggesting both a high level of grievance in society and a "stolen" election. -->

## Election manipulation and incumbent strength

Finally, an alternative theory of election manipulation challenges the assumption that an incumbent's ability manufacture votes communicates nothing about its ability to defeat a protest movement. This view, most persuasively articulated by Simpser [-@simpser2013], holds that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement. Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors [@simpser2013]. Since this view holds that more manipulation can further entrench the incumbent, it is a normatively pessimistic one.

We can thus consider four possible states of the world, based on which of the foregoing assumptions are true in fact: one where both assumptions hold, one where neither holds, and two where only one assumption holds. The normatively optimistic view of the relationship between election manipulation and protest requires both assumptions to be true: electoral manipulation must communicate little information about incumbents' ability to outlast protest, and protest initiation must be narrowly strategic. If either one of these assumptions fails to hold, the deterrence logic is undermined. However, empirically distinguishing between these four possible states of the world is further complicated by the fact that work on this subject, whether empirical or purely formal, has largely utilized a binary measure of protest occurrence as the outcome of interest. Unfortunately, as Table 1 shows, three of four possible states of the world make the same prediction: that increasing manipulation will result in an increased likelihood of protest.

```{=latex}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\begin{tabular}{@{}rll@{}}
\toprule
Incumbent's genuine popularity:                & \multicolumn{2}{c}{Low}                                               \\ \midrule
\multicolumn{1}{r|}{Protest is assumed to be:} & \multicolumn{1}{c|}{Instrumental} & \multicolumn{1}{c|}{Multipurpose} \\ \midrule
\multicolumn{1}{|l|}{Manipulation...}          &                                   & \multicolumn{1}{l|}{}             \\ \midrule
\multicolumn{1}{|r|}{...obscures strength / indicates weakness} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Luo and Rozenas 2020; \\ Chernykh and Svolik 2015; \\ Little, Tucker, and LaGatta 2015)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007; \\ Beaulieu 2014; \\ Brancati 2016)\end{tabular}} \\ \midrule
\multicolumn{1}{r|}{...reflects strength} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and decreasing\\ \\ (Simpser 2013)\end{tabular}} &
  \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}High and increasing\\ \\ (Tucker 2007;\\ Beaulieu 2014;\\ Brancati 2016)\end{tabular}} \\ \bottomrule
\end{tabular}
\caption{Likelihood of protest initiation and expected relationship with increasing election manipulation}
\label{tab:initiation}
\end{table}
```

This ambiguity may help account for the inconsistent relationship between election manipulation and protest found in the literature. Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020;@Hyde2014]. A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012]. Similar results are found by Rød [-@Rod2019], using a different measure of manipulation.  Using an alternative measure of manipulation from the NELDA dataset, however, Hafner-Burton et al [-@Hafner-Burton2016] find that the association between election fraud and post-election protest falls short of conventional statistical significance. Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022]. As noted earlier, Simpser finds a negative relationship between the severity of election manipulation and the likelihood of protest [@simpser2013] and a positive relationship between 'excessive manipulation' (i.e. extensive manipulation and a large incumbent vote-share) and regime tenure.

Because this paper attempts to systematize and test two competing models of electoral manipulation and protest, it is worth spelling out each expectation and its mechanisms in detail. To do so, I will describe each theory's expectations under two worlds: one where the incumbent's genuine popularity is low, and one where it is high. For each of these notional worlds, expectations for protest initiation and duration are explored under both sets of assumptions about the nature of protest and electoral manipulation.

First, consider the a world in which the incumbent's true popularity is low; this is presumed to be the case for most authoritarian regimes by their nature. Next, let us assume that the regime-weakness model is correct: election manipulation indicates weakness (or obscures information about strength). Finally, assume that protest is elite-led and instrumental. This is the world of the classic regime-weakness framework: we should expect to see protest initiation become more likely as the severity of electoral manipulation increases.

The same prediction holds if we assume that protest is largely multi-purpose, while maintaining the assumptions that true popularity is low and manipulation indicates weakness. This is the classic focal point example [@tucker2007]. Here, the incumbent's low genuine popularity implies high levels of grievance among the population, which are increasingly likely to be focused into mass protest as manipulation increases (i.e. as the election becomes more likely to appear 'stolen').

Now, while maintaining the assumption that the incumbent's popularity is low, let us assume that increasing electoral manipulation indicates an otherwise stronger incumbent. In this case, if protest is primarily instrumental, we should expect opposition groups to shy away from protest initiation as electoral manipulation increases. An unpopular incumbent who manipulates extensively is stronger, in this framework, than an unpopular incumbent who does not. As a result, we should expect the opposite prediction of the classic regime-weakness model.

However, if the incumbent's popularity is low, manipulation indicates underlying strength, and protest is largely dispersed and multi-purpose, the expectation is reversed. Just as in the regime-weakness case, low popularity indicates high grievances. Increasing electoral manipulation helps focus those grievances and drives protest initiation.

The key takeaway, then, is that for most authoritarian regimes (where genuine incumbent popularity is likely low), an examination of protest initiation alone cannot distinguish between three out of four possible data-generating processes. A pattern in which higher levels of manipulation are associated with increasing risk of protest initiation can be explained either by the regime-weakness model, or by a world in which manipulation indicates incumbent staying-power but protest is non-instrumental.

@@

How, then, would this view of electoral manipulation map onto the instrumental vs. multipurpose views of protest discussed above? While regime-weakness models have exclusively (to my knowledge) been framed around protest initiation, it is straightforward to extend their predictions to protest duration. Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods. For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the less likely the incumbent is to survive a challenge. In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge. Increasing manipulation indicates deeper unpopularity (holding vote-share constant), so we should expect a less popular incumbent to fail more quickly. The same logic holds, with slightly different mechanisms, if protests are assumed to be multipurpose. Here, greater manipulation is used to overcome deeper unpopularity (i.e. widespread grievances); more manipulation indicates an incumbent facing an increasingly powerful challenger. Thus, the regime-weakness model of manipulation should predict that the duration of unsuccessful protests will be increasing in election manipulation: more manipulation indicates a stronger challenger, which will take longer to defeat on average.

@@

Need to discuss vote-share next:

> H1: Increasing election fraud makes protest more likely, with the size of this effect decreasing with incumbent-party vote-share.

> H2: Election manipulation is positively associated with protest movement duration, with the size of this effect decreasing with incumbent vote-share.

@@

Testing H1 and H2 together will help to evaluate the regime-weakness model of electoral manipulation. If H1 is supported and H2 is not, it would provide evidence in favor of the regime-weakness model; manipulation would appear associated with incumbent weakness, and protest initiation would appear to take advantage of such. On the other hand, if both hypotheses are supported, it would endorse the argument that electoral manipulation can be associated with sources of incumbent strength in a world where the aims of electoral protest are broadly (rather than narrowly) strategic.

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.33, 2.2]")) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

>>>>

---
output:
  pdf_document:
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
abstract: "Two models of the relationship between election manipulation and protest can be found in the literature. One approach holds that election fraud reveals an incumbent who is more vulnerable to protest; in this view, protest risk can result in cleaner elections. An alternative approach contends that election manipulation can be a sign of underlying regime strengths; under this framework, protest risk is less likely to deter manipulation. This article interrogates the assumptions underlying these models, and conducts a novel test of their implications. Using data on the duration of protest waves in 659 authoritarian cases from 1990-2012, this paper shows that governments which manipulate elections more also defeat protests more quickly. This supports the ‘regime-strength’ model, and helps explain why electoral authoritarianism is a sticky equilibrium amid ongoing democratic backsliding. It suggests researchers should look beyond protest risk in models of electoral manipulation."
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
  - \usepackage{multirow}
  - \usepackage{lscape}
  - \usepackage{pdflscape}
  - \usepackage{longtable}

@@

At first glance, it would appear straightforward to test these competing models. The regime-weakness model typically predicts that increasing manipulation (at a constant level of true popularity) increases the likelihood that opposition groups will initiate post-election protest [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a]. A low total vote-share amid severe manipulation is the strongest possible signal of incumbent weakness; it has had to rely on the 'cheap talk' of manipulation to produce paltry results, indicating that the incumbent's genuine popularity is very low and its likelihood of defeating the protest movement is likewise small. By contrast, the regime-strength model would yield the opposite prediction [@simpser2013]; that protest initiation decreases with the level of manipulation (holding vote-share constant). Under the regime-strength model, the most favorable  election outcome for protesters is one where the incumbent's vote-share and manipulation level are low, a combination which indicate low natural popularity and diminished resource capacity.

@@

As sketched briefly above, a fruitful literature has examined the connections between election integrity and mass protest. This literature has grappled with the extent to which election manipulation reflects, and reveals, the incumbent's strength. Various authors have put forth different modalities of incumbent strength, with some focusing on popularity and others on coercive power. This paper takes a general approach, following Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025]: a stronger incumbent is one who is better able to survive in power throughout the whole electoral cycle. This means the incumbent has both prevailed in the election itself, and managed to hold off any subsequent popular revolts or elite challenges [@svolik_politics_2012]. As Leventoğlu et al [-@leventogluSynthesizingTheoriesAuthoritarian2025] discuss in their excellent analysis, genuine popularity is but one potential source of incumbent strength. Popularity certainly helps an incumbent win an election and survive any challenges in its aftermath, but it is only one tool; incumbents may also draw on resource wealth, organizational capacity, support among influential groups, loyal military and security services, or help from foreign patrons. These latter factors may be weakly correlated, if at all, with the incumbent's genuine popularity.

@@

@@

The characteristic assumption in regime-weakness models is that electoral manipulation communicates nothing about incumbent strength, which is closely identified with popularity. Since more manufactured votes indicate deeper unpopularity, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent. For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p. 1153). In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud---a pro-democracy outcome in which stronger opposition groups deter manipulation by incumbents. Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010].

@@

This assumption, however, is contested---most prominently by Simpser [-@simpser2013]---who argues that increasing levels of election manipulation can indicate a well-resourced, well-organized, powerful incumbent (independent of its popularity); these resources and capacities, in turn, may help the incumbent weather a protest movement. Under this view, extensive election manipulation can be a show of strength, which induces cooperation with the incumbent from other societal actors [@simpser2013].

Prior research on this subject has typically treated protest initiation as the dependent variable. This approach is limited, however, since both the regime-weakness model and Simpser's alternative view can be made to make the same predictions for protest initiation--namely that increasing manipulation makes protest more likely. The key is in the second assumption of the regime-weakness framework, which holds that protest initiation will occur only when opponents perceive the incumbent as weak. A broader view of protest, well-supported in the literature, indicates that protests can emerge when manipulation is severe, even when incumbents are strong.

@@

Regime-weakness models assume that opposition protest is *elite-led and instrumental*. The opposition is typically modeled as a unitary, strategic actor who stages protest when the expected benefits of successfully overthrowing the incumbent outweighs the expected cost of failure. The signal of weakness sent by election manipulation, in turn, indicates to opposition actors that protest is more likely to succeed in unseating the incumbent [@Luo2017;@Little2012;@Schedler2002;@Beaulieu2014a]. Incumbents in these models typically face a tradeoff between the short-term benefits of electoral manipulation for winning the election, and any weakness that relying on fraud reveals to the opposition [@Rozenas2016b]. As Beaulieu [-@Beaulieu2014a, p. 48] writes in this tradition, "As the opposition grows stronger, the amount of [manipulation] he will tolerate shrinks, and the incumbent who wishes to avoid protest will need to reduce the amount of manipulation she commits." Such abstraction helps keep formal models tractable, but this assumption is difficult to square with recent research on electoral protest and a broader understanding of political protest in authoritarian regimes.

Electoral protest may occur because manipulation galvanizes citizens who hold diverse grievances [@tucker2007], because economic conditions are poor [@Brancati2016;@Beaulieu2014a], because electoral manipulation makes people angry [@Szakonyi2021], or simply because their party lost the election [@andersonLearningLoseElection2006]. Opposition elites, in turn, might seek to initiate---or amplify---protest to take advantage of this popular discontent for broad, rather than instrumental, reasons. They may wish to pressure the incumbent in order to push for improved competitiveness in the *next* election or to bargain for policy concessions [@chernykh2014;@robertson_protest_2010]. They may stage protests in order to identify new participants, forge new social ties [@barberaModelProtestsRevolution2019] and test new tactics. They may seek to communicate information about the government's unpopularity, to set the stage for larger, future protests [@lohmannDynamicsInformationalCascades1994]. Protests may be staged as part of internal factional rivalries within the opposition [@Buckles2017], or to drive media coverage of issues important to the opposition [@huangMediaProtestDiffusion2019;@jenningsStreetDemonstrationsMedia2019]. Indeed, opposition elites may take advantage of the bare fact of an election to challenge the incumbent, irrespective of the level of manipulation or the incumbent's vote-share [@Pop-Eleches2015].

@@

The varied motivations for protest may help account for the inconsistent relationship between election manipulation and protest initiation found in the literature. Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Harvey2020;@Hyde2014]. A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012]. Similar results are found by Rød [-@Rod2019], using a different measure of manipulation.  Using an alternative measure of manipulation from the NELDA dataset, however, Hafner-Burton et al [-@Hafner-Burton2016] find that the association between election fraud and post-election protest falls short of conventional statistical significance. Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022]. As noted earlier, Simpser finds a negative relationship between the severity of election manipulation and the likelihood of protest [@simpser2013] and a positive relationship between 'excessive manipulation' (i.e. extensive manipulation and a large incumbent vote-share) and regime tenure.

@@

How, then, would this view of electoral manipulation map onto the instrumental vs. multipurpose views of protest discussed above? While regime-weakness models have exclusively (to my knowledge) been framed around protest initiation, it is straightforward to extend their predictions to protest duration. Regime-weakness models of electoral manipulation treat election manipulation and genuine votes as substitute goods. For any given vote-share, an increase in manufactured votes reduces the number of genuine votes; that is, the higher the number of manufactured votes for any given vote total, the lower the incumbent's genuine popularity is and the less likely they are to survive a challenge. In an instrumental mode of protest, strategic opposition actors perceive the incumbent's relative lack of popularity, and count on the majority of the public to back their challenge. Increasing manipulation indicates deeper unpopularity (holding vote-share constant), so we should expect a less popular incumbent to fail more quickly. The same logic holds, with slightly different mechanisms, if protests are assumed to be multipurpose. Here, greater manipulation is used to overcome deeper unpopularity (i.e. widespread grievances); more manipulation indicates an incumbent facing an increasingly powerful challenger. Thus, the regime-weakness model of manipulation should predict that the duration of unsuccessful protests will be increasing in election manipulation: more manipulation indicates a stronger challenger, which will take longer to defeat on average.

@@

## Empirical implications: vote-share vs. genuine popularity

Both models of election manipulation consider genuine popularity to be a source of regime strength, and assume that increasing electoral manipulation either distorts or weakens this indicator by mingling genuine votes with manufactured ones. The distinction between the two models is over whether or not manipulation simultaneously indicates additional sources of strength other than popularity.

Of course, in most authoritarian cases, the incumbent's true popularity is unknown due to censorship, repression of opposition candidates, and preference falsification by citizens. As a result, scholars (and political actors in real time) are left to use the incumbent's vote-share and perceived level of manipulation as imperfect indicators of the incumbent's true popularity. Since genuine popularity is indicated by the mix of incumbent vote-share and level of manipulation, both indicators must be included in our hypotheses. Table \@ref(tab:durations) summarizes how both models of electoral manipulation characterize incumbent strength under varying conditions, each of which is laid out in more detail below.

```{=latex}
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{multirow}
\begin{table}[]
\begin{tabular}{@{}lllll@{}}
 &  & \multicolumn{2}{c}{Manipulation} & Duration prediction of defeated protests \\
 &  & Low            & High            &                                          \\
\multicolumn{1}{c}{\multirow{2}{*}{Vote-share}} &
  High &
  \begin{tabular}[c]{@{}l@{}}Popularity: High\\ MS: Mixed\\ \\ \\ Predicted strength (RW): Strong\\ Predicted strength (RS): Unclear\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Popularity: Low (RW) or mixed (RS)\\ MS: High\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Moderate\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: No significant effect with manipulation\end{tabular} \\
\multicolumn{1}{c}{} &
  Low &
  \begin{tabular}[c]{@{}l@{}}Popularity: Low\\ MS: Low\\ \\ \\ Predicted strength (RW): Weak\\ Predicted strength (RS): Weak\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}Popularity: Very low\\ MS: High\\ \\ \\ Predicted strength (RW): Very weak\\ Predicted strength (RS): Moderate\end{tabular} &
  \begin{tabular}[c]{@{}l@{}}RW: Duration increases with manipulation\\ RS: Duration decreases with manipulation\end{tabular}
\end{tabular}
\caption{Predicted incumbent strength, and associated hypotheses for protest duration, under regime-weakness (RW) and regime-strength (RS) models of manipulation as incumbent vote-share and level of electoral manipulation vary. 'MS' indicates incumbent's underlying strength associated with manipulation under regime-strength model (i.e. 'manipulation strength'). Duration predictions for protests which the government ultimately defeats.}
\label{tab:durations}
\end{table}
```

The upper left cell in Table \@ref(tab:durations) illustrates a scenario where the incumbent's vote-share is high and manipulation is low. This combination indicates a relatively popular incumbent, meaning it should be characterized as 'strong' under the regime-weakness model of manipulation. However, cases in this cell are less informative about 'manipulation strength,' those capacities and resources that electoral manipulation draws on. This cell could include regimes that lack the capacity to manipulation, *or* regimes which have such capacities but choose not to employ them in a particular election. As a result, from the regime-strength perspective, this cell would likely include a mix of weaker and stronger regimes. There is thus no clear expected tendency for this group under that model.

Moving to the right in Table \@ref(tab:durations), incumbent vote-share remains high but manipulation is also elevated. For the regime-weakness model, cases in this cell are clearly expected to be weaker than those in the top left; one's high vote-share mostly reflects genuine popularity, while the other's is largely manufactured. In the regime-weakness model, regimes should unambiguously become weaker as we move to the right in this row. As a result, holding vote-share contant at a high level, increasing manipulation should lead to longer-lasting protests under the regime-weakness model.

A different prediction emerges from the regime-strength framework. Here, too, increasing manipulation may indicate broadly lower popularity (holding vote-share constant). But it may also indicate a popular incumbent relying on manipulation tools to boost turnout in a low-competition election [@Smyth2018;@martinezicomaElectionTurnoutAuthoritarian2020]. In any event, a decrease in genuine popularity is expected to be offset by increased strength indicated by the ability to manipulate the election severely. As a result, overall strength is expected to be moderate. Where the regime-weakness model predicts a sharp decline in incumbent strength moving from the top left to the top right of Table \@ref(tab:durations), the regime-strength model suggests there should be no consistent effect on protest duration as manipulation increases, since decreases in popularity are offset by increases in other sources of power.

Moving to the lower left, the incumbent's vote-share is now assumed to be low, as is the level of electoral manipulation. Both models of manipulation agree here: such incumbents are weak. They lack both genuine popularity, and the ability to produce significant manipulation. Under the regime-strength framework, such an incumbent would have had strong incentives to produce more manipulation if he could have; the low level of manipulation is thus an indicator of low capacity.

Finally, moving to the lower right, the level of manipulation increases while vote-share remains low. This is the weakest of all incumbents in the regime-weakness framework. Holding vote-share constant at a low value, increasing manipulation indicates the incumbent's genuine popularity---and thus strength---is exceedingly low. For the regime-strength model, cases in this cell are expected to be stronger than those to the left, as their increasing capacity to manipulate elections indicates sources of strength beyond their (declining) popularity.

Moving left-to-right along Table \@ref(tab:durations) shows two clear hypotheses; one direct hypothesis under the regime-weakness model and one interaction effect under the regime-strength model. First, the regime-weakness model should predict that protest duration will increase with manipulation when incumbent at any vote-share (as 'cheap-talk' manufactured votes displace genuine popularity in the vote-share). The first hypothesis is straightforward.

>H1 (Regime-weakness): Increases in electoral manipulation will be associated with increasing duration of unsuccessful electoral protest.

Second, the regime-strength model predicts that protest duration will *decrease* as electoral manipulation increases, but only when the incumbent's vote-share is relatively low. At higher vote-shares, the model predicts ambiguous relationships between vote-share, manipulation, and overall regime strength, suggesting that statistical correlations should weaken.

>H2 (Regime-weakness): Increases in electoral manipulation will be associated with decreasing duration of unsuccessful electoral protest when incumbent vote-share is low.

@@

Testing H1 and H2 together will help to evaluate the regime-weakness model of electoral manipulation. It is important to note that the theories make contradictory predictions. H1 predicts an unconditional *positive* relationship between manipulation and protest duration, while H2 predicts a *negative* relationship, conditional on vote-share.

[Note that the type of protest that happens (instrumental vs. multipurpose) will be examined in the discussion / appendix. Consider also coming up with names for 'manipulative strength']

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag +  + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.29, 2.2]")) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()
